Seismic Hazard Estimation of Northern Iran Using Smoothed Historical Seismicity

### 2016-01-11 analysis_parameters.tex
Region of interest and Seismic hazard are completed.

\newpage
\section{Analysis Parameters}
\subsection{Magnitude Interval}

In this study we use two sets of data, i.e., historical and instrumental seismic catalogs. We use the historical catalog which is converted into moment magnitude and reported by \citep{Karimiparidari2013}. Fig.~\ref{fig:historical} shows the historical earthquakes (pre-1900) in northern Iran. The instrumental data has been reported mainly by the International Institute of Earthquake Engineering and Seismology of Iran \citep{IIEES}, and are reported in $M_L$ magnitude type. In this study we use the relationship that proposed by \citet{Karimiparidari2013}, to convert local magnitude into moment magnitude.

\begin{figure} [H]
\centering
\includegraphics[scale=0.8]{figures/pdf/Figure2.pdf} 
\caption{Historical earthquakes of Iran (pre 1900). Different colors represent different year of occurrences and size of circles are proportional to the earthquake magnitude. }
\label{fig:historical}
\end{figure}

\subsection{Minimum Magnitude}
\subsection{Attenuation Relationship}
The choice of a ground motion attenuation model is of great importance since attenuation has proven to be a highly influential factor of seismic hazard. \citet{Zafarani2014} used 163 free-field acceleration time histories recorded at epicentral distance of up to 200 km from 32 earthquakes to investigate the predictive capabilities of the local, regional, and next generation attenuation (NGA) ground-motion prediction equations and determined their applicability for northern Iran. After evaluating different Ground Motion Prediction Equations, \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008} represented suitable performance for  PGA  with LLH (Log-Likelihood method) score of 1.54, 1.55, and 1.59. Mean of the mentioned attenuation relationships (not shown here), are very close together, especially at higher magnitude. Using \citet{Scherbaum2009} approach and \citet{Zafarani2014} coefficients we calculated the logic tree weights as 0.3376, 0.3354, and 0.3270, respectively. Getting close mean values and logic tree weights, in order to consider the epistemic uncertainty, instead of using several GMPEs we use the top rank GMPE (i.e.  \citet{Kalkan2004} ) with  $\pm$ standard deviation with weight of 0.2 (for each of standard deviation branches) and 0.6 for the mean value.    \\

@@

\begin{table}[h]
\centering
\caption{Seismicity parameters for three tectonic seismic regions.}
    \begin{tabular}{ccccc}
    ~                   & $b-value$            & $\lambda$                  & $M_{max}$ Calculated & $M_{max}$ obseved \\ \hline
    Azerbaijan    & 1.06 $\pm$ 0.03  & (1.827  $\pm$ 0.216) & 7.78  $\pm$ 0.26           & 7.7          \\ \hline
    Alborz           & 1.1  $\pm$ 0.03   & (2.290  $\pm$ 0.282) & 7.91  $\pm$ 0.27           & 7.7          \\ \hline
    Kopeh Dagh & 0.97  $\pm$ 0.04 & (1.716  $\pm$ 0.243) & 7.67  $\pm$ 0.26           & 7.6          \\
    \end{tabular}

\end{table}

>>>>

\newpage
\section{Analysis Parameters}
\subsection{Magnitude Interval}
\subsection{Minimum Magnitude}
\subsection{Attenuation Relationship}
The choice of a ground motion attenuation model is of great importance since attenuation has proven to be a highly influential factor of seismic hazard. \citet{Zafarani2014} used 163 free-field acceleration time histories recorded at epicentral distance of up to 200 km from 32 earthquakes to investigate the predictive capabilities of the local, regional, and next generation attenuation (NGA) ground-motion prediction equations and determined their applicability for northern Iran. After evaluating different Ground Motion Prediction Equations, \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008} represented suitable performance for  PGA  with LLH (Log-Likelihood method) score of 1.54, 1.55, and 1.59. Mean of the mentioned attenuation relationships (not shown here), are very close together, especially at higher magnitude. Using \citet{Scherbaum2009} approach and \citet{Zafarani2014} coefficients we calculated the logic tree weights as 0.3376, 0.3354, and 0.3270, respectively. Getting close mean values and logic tree weights, in order to consider the epistemic uncertainty, instead of using several GMPEs we use the top rank GMPE (i.e.  \citet{Kalkan2004} ) with  $\pm$ standard deviation with weight of 0.2 (for each of standard deviation branches) and 0.6 for the mean value.    \\

@@

\begin{table*}[!ht]
\centering
\caption{Seismicity parameters for three tectonic seismic regions.}
    \begin{tabular}{ccccc}
    ~                   & $b-value$            & $\lambda$                  & $M_{max}$ Calculated & $M_{max}$ obseved \\ \hline
    Azerbaijan    & 1.06 $\pm$ 0.03  & (1.827  $\pm$ 0.216) & 7.78  $\pm$ 0.26           & 7.7          \\ \hline
    Alborz           & 1.1  $\pm$ 0.03   & (2.290  $\pm$ 0.282) & 7.91  $\pm$ 0.27           & 7.7          \\ \hline
    Kopeh Dagh & 0.97  $\pm$ 0.04 & (1.716  $\pm$ 0.243) & 7.67  $\pm$ 0.26           & 7.6          \\
    \end{tabular}

\end{table*}

### 2016-01-11 seismic_hazard.tex
adding new section files, journal of seismology format's file.
Region of interest and Seismic hazard are completed.

\section{Methodology for Hazard calculation}

To calculate the seismic hazard, we use spatially smoothed seismic hazard analysis \citep{Frankel1995}. In this model, seismic events are spatially gridded to cells. We are attempting to assess the relative likelihood of moderate earthquakes ($M_w > 4.5$), which cause structural damage. According to \citet{Frankel1995}, moderate earthquakes generally occur in areas that there have been significant numbers of events of magnitude 3 and above. Therefore, these events provide a reasonable guide to where moderate earthquakes will most likely occur. Since the catalog's completeness for $M3+$ and $M4+$ is different, we use two magnitude completeness range for less than $M_w5$ . Category 5+ and others (M6+ and M7+) assume that future $M_w4.5$ events will occur near where they have occurred in the past. \\
\noindent
According to Building and Housing Research Center  \citep{BHRC2014}, $M_w5$ is considered as a threshold magnitude to structural damage. In the smoothed seismicity method, the model uses each event location as a point source, hence $M_w4.5$ could be damaging earthquake if it occurs very close to the structure. In this study, in order to consider the probabilistic seismic hazard of both models, we defined two models based on $M_w5$ and $M_w4.5$ as a threshold magnitude for seismic hazard calculation. \\
\noindent
\noindent
In order to calculate the annual rate ($\lambda$) of exceeding ground motion (u) at a specific site, according to \citet{Frankel1995}, first we need to divide the region into cells (0.1*0.1) and count all earthquake events which are bigger than $M_{ref}$ in each cell. We use Herrmann formula \citep{Herrmann1977} to convert the cumulative (i.e. number of events bigger than $M_{ref}$) values to incremental (i.e. number of events from$M_{ref}$ to $M_{ref}+\Delta_m$) values. Using Gaussian function we smooth the number of events in each cell. The smoothed value $\tilde{n}_i$ is obtained from

@@

>>>>

\section{Seismic Hazard}

\subsection{Background}

Regarding historical earthquakes and the need for constructing new buildings and infrastructure, different seismic hazard analyses were conducted in Iran. Dividing Iran into 20 tectonic seismic regions, \citet{Tavakoli1999} prepared iso-acceleration contour lines and seismic hazard zonation for return periods of 75 and 475 years or 50\% and 10\% probability of exceedance in 50 years, based on major known lines and area source models. According to their study, North Tabriz fault zone and north of Tehran fault zone are in the highest acceleration contour. Maximum mean acceleration in the vicinity of these tectonic elements is predicted to be around 0.45 $g$ and 0.3 $g$ for a return period of 475 and 75 years, respectively. In the smaller scale, using the logic tree approach to compensate for uncertainties in the attenuation relationship, \citet{Ghodrati2003} presented a probabilistic seismic hazard assessment of metropolitan Tehran, the capital of Iran. The results showed that the PGA ranges from 0.27-0.46 $g$ for a return period of 475 years and from 0.33-0.55 $g$ for a return period of 950 years. 
In this study we use background seismicity to calculate the seismic hazard.Using background seismicity, \citet{Cao1996} conducted a seismic hazard estimation study in southern California. They concluded that one could obtain similar results implementing background seismicity to those obtained by introducing zones for areas with well-recorded seismicity patterns. \citet{Lapajne1997} used spatially smoothed seismicity modeling to acquire a seismic hazard outlook in Slovenia. They defined 4 different models with different correlation distances, including a model based of the total released seismic energy. They gained an acceptable PGA in comparison with other seismic hazard approaches. \citet{Akinci2004} estimate the seismic hazard in central and northern Italy using smoothed historical seismicity. Their results showed that the smoothed seismicity approach gives reasonable regionalized results without introducing seismogenic zones. This approach has also been used in many regions with different seismicity patterns \citep{Wesson1999, Klein2001, Hamdache2008, Kalkan2009, Moschetti2014, Boyd2008}.

\subsection{Analysis Method}
\subsubsection{smoothed seismicity method}
To calculate the seismic hazard, we use spatially smoothed seismic hazard analysis \citep{Frankel1995}. In this model, seismic events are spatially gridded to cells. We are attempting to assess the relative likelihood of moderate earthquakes ($M_w > 4.5$), which cause structural damage. According to \citet{Frankel1995}, moderate earthquakes generally occur in areas that there have been significant numbers of events of magnitude 3 and above. Therefore, these events provide a reasonable guide to where moderate earthquakes will most likely occur. Since the catalog's completeness for $M3+$ and $M4+$ is different, we use two magnitude completeness range for less than $M_ w5$ . Category 5+ and others ($M6+$ and $M7+$) assume that future $M_w 4.5$ events will occur near where they have occurred in the past. \\
\noindent
According to Building and Housing Research Center  \citep{BHRC2014}, $M_w5$ is considered as a threshold magnitude to structural damage. In the smoothed seismicity method, the model uses each event location as a point source, hence $M_w4.5$ could be damaging earthquake for old masonry buildings or even engineering building if it is very close to it, in this study, in order to consider the probabilistic seismic hazard of both models, we defined two models based on $M_w 5$ and $M_w 4.5$ as a threshold magnitude for seismic hazard calculation. \\
\noindent
\noindent
In order to calculate the annual rate ($\lambda$) of exceeding ground motion ($u$) at a specific site, according to \citet{Frankel1995}, first we need to divide the region into cells (0.1*0.1) and count all earthquake events which are bigger than $M_{ref}$ in each cell. We use Herrmann formula \citep{Herrmann1977} to convert the cumulative (i.e. number of events bigger than $M_{ref}$) values to incremental (i.e. number of events from$M_{ref}$ to $M_{ref}+\Delta_m$) values. Using Gaussian function we smooth the number of events in each cell. The smoothed value $\tilde{n}_i$ is obtained from

@@

\subsubsection{ingredients}
